---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

``` {r}
MyData <- read.csv(file="activity.csv", header=TRUE, sep=",")
head(MyData)
tail(MyData)
summary(MyData)
```
## What is mean total number of steps taken per day?
####1. Make a histogram of the total number of steps taken each day

```{r hist, fig.width=8, fig.height=6}
numberofstepsperday <- aggregate( formula = steps~date, data = MyData, FUN = sum ) 
hist(numberofstepsperday$steps, main="Histogram of Steps Taken Each Day",
      xlab="Steps Taken Per Day", col="red")
```

####2. Calculate and report the mean and median total number of steps taken per day
```{r}
meanData <- aggregate( formula = steps~date, data = MyData, FUN = mean )
meanData
medianData <- aggregate( formula = steps~date, data = MyData, FUN = mean )
medianData
```
## What is the average daily activity pattern?

####1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r timeseries, fig.width=8, fig.height=6}
meanData5minInterval <- aggregate( formula = steps~interval, data = MyData, FUN = mean )
meanData5minInterval 
plot(x=meanData5minInterval$interval, y=meanData5minInterval$steps, type="l", 
      main="Average Number of Steps Taken Each Day for Each 5 min Interval",
      xlab="5 Min Interval", ylab="Average Number of Steps Taken", col="red")

```

####2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxstepatwhich5mininterval <- meanData5minInterval$interval[meanData5minInterval[,2]==max(meanData5minInterval$steps)]
maxstepatwhich5mininterval
```
## Imputing missing values
```{r}
# Replace NA steps with 0
MyData[is.na(MyData)] <- 0
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
DayofWeek <- c("Weekend", "Weekday", "Weekday", "Weekday", "Weekday", 
               "Weekday", "Weekend")[as.POSIXlt(MyData$date)$wday + 1]
MyNewData <- cbind(MyData,DayofWeek)
summary(MyNewData)

weekendData <- aggregate( formula = steps~interval, data = subset(MyNewData,DayofWeek == "Weekend", select=c(1:4)), FUN = mean )

weekdayData <- aggregate( formula = steps~interval, data = subset(MyNewData,DayofWeek == "Weekday", select=c(1:4)), FUN = mean )

plot(x=weekendData$interval, y=weekendData$steps, type="l", 
      main="Average Number of Steps Taken Each Day for Each 5 min Interval",
      xlab="5 Min Interval", ylab="Average Number of Steps Taken", ylim=c(0,250), col="red")
lines(x=weekdayData$interval, y=weekdayData$steps,col="blue")
legend(x="topleft", c("Weekend","Weekday"),
lty=c(1,1), # gives the legend appropriate symbols (lines)
lwd=c(2.5,2.5),
col=c("red","blue")) # gives the legend lines the correct color and width


```

